Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-27 drafts/hybrid_zone.tex
edits, mostly to discussion

@@

The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact 
before reproductive isolation is complete, gene flow is still possible. 
As two such populations interbreed,
individuals' genomes can be labeled according to which ancestral population they are inherited from,
and local interbreeding and migration of such populations
creates a gradient of ancestry assignments across geography,
centered on their point of contact \citep{Barton1985}. 
\plr{(check that neutral clines discussed in barton '85)}
If the populations are sufficiently diverged,
this process leaves a distinct pattern of variation across the genome, 
in which long tracts of divergent haplotypes from each ancestral population  
are broken up by historical recombination events,
leading to substantial correlations in ancestry between distant markers
(termed ``ancestry LD'').
\plr{(but, ``ancestry linkage disequilibrium'' doesn't work;
how about just ``correlations in ancestry''?}

@@

Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.  
\yb{The theoretical expectations developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}

\plr{explain relationship to \citet{Sedghifar2015}}

@@

\paragraph{Establishment of the cline at the selected locus}
The deterministic theory predicts that after secondary contact,
the two alleles at the selected locus will form a stable cline \citep{Barton1979},
affecting neighboring loci as well.
First, we turn our attention to how the cline at the selected locus itself is formed.
Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
Suppose that secondary contact occurred at time $t=0$, 
and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
% (In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
Let $\sigma^2$ be the mean squared distance between parent and offspring 
(i.e., the variance of the distribution of dispersal distance).
Assuming that: (i) alleles locally assort into diploids randomly, 
(ii) habitat and dispersal are homogeneous, and 
(iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
\begin{align} \label{eqn:cline_pde}
    \dot p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
\end{align}
where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
as derived in \citet{Bazykin1969}; for discussion of diploidy see \citep{christiansen1995genotypic}.
Although there is no known exact analytic solution to this equation, even in steady state, 
a good approximation to the equilibrium solution is that
$\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}. 
This relies on numerous approximations, but the general conclusions should be quite robust: 
the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} results in a dimensionless equation,
implying that the cline establishes over a timescale of the order $1/s$. 
In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes. 
\alisa{I don't think I understand the independence from $\sigma$?}
\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}

@@

At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
Therefore, roughly, lineages on ``their own'' side wander randomly,
while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
back towards the side where they are more common. 
\alisa{Is this a construction that we've enforced, or should this maybe go into the results section?}
\plr{This isn't a construction, it's a consequence of the model, so you're right, it is maybe a "result", but I think the explanation of why it happens belongs here.}
Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
Lineages of $B$ alleles do the same thing, of course, but in the opposite direction. 
% \alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}

@@

\begin{figure}
    \includegraphics{figs/linked-frequencies}
    \caption{
        \textbf{Probabilities of $A$ ancestry,}
        across space (vertical axis, in units of $\sigma$) 
        and time (horizontal axis, in generations).
        In each plot, color corresponds to the expected frequency of $A$ ancestry
        at a particular location in time and space.
        The selection coefficient is $s=.02$.
        \textbf{Top left:} at the selected site, showing establishment and stabilization of the cline
        on a time scale of $1/s=50$ generations.
        \textbf{Bottom left:} at an unliked site,
        with cline flattening continuing with $\sqrt{t}$.
        Remaining figures show frequencies of $A$ ancestry \emph{conditional}
        on the ancestry at the selected site,
        at different distances from the selected site ($r=.01$, .04, and 0.5 Morgans),
        as described in the text (see definition of $q_z(x,t,r)$).
        See figure \ref{sfig:linked_cline_heatmaps_longtime}
        for the same figure over a longer period of time.
    }
    \label{fig:linked_cline_heatmaps}
\end{figure}

\paragraph{Linked clines}
We can use this diffusion model for lineages to compute expected clines in ancestry,
i.e., the expected proportion of individuals who inherit from species $A$,
as a function of space, time, and position on the genome.
Precisely, we need the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$,
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
We denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
(Recall that $Z$ depends implicitly on $r$.)
The description above implies that $q_z$ solves the following Kolmogorov backward equation:
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            r p(x,t) (q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} 
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
applied to a backwards-time process!}

@@

Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1$ and $s=0.01$. From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$). These were compared to the output of neutral simulations with $s=0$
\plr{missing end of sentence?}

The simulations are executed in R, with scripts available on
\url{https://github.com/asedghifar/hybridzones}.

@@

Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.

Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$, that flank the block of unbroken ancestry containing $m$.

@@

@@

\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
    which we have done using both theory and simulation.
    Our most useful observations are that clines establish over time $1/s$,
    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
    tend to move back towards the side where they are common
    at speed $\sigma/\sqrt{s}$.
    This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
}

Most attempts to identify loci under selection in hybrid zones do so by identifying exceptional allele frequencies as a function of geographic position or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
Current genomic approaches \alisa{applied} to the study of hybrid zones provide a additional rich sources of information -- namely ancestry deconvolution facilitated by programs such as hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome.  \alisa{In particular,} the information contained in the length distribution of these ancestry tracts genome-wide  has been extensively used to infer the demographic history of hybridization and admixture \citep[e.g. ][]{Hellenthal2014}.  \alisa{However,the impact of selection on ancestry tract lengths in hybrid zones has received much less attention. Here, we present an expectation of block-length patterns throughout a hybrid zone, and find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.}

@@

\plr{(selected lineages drift back, so haplotypes are longer)}
Our finding that long blocks of ancestry surround selected sites in hybrid zones reflects the intuition that regions surrounding selected loci are going to \alisa{introgress less readily?} have a harder time introgressing. 
Thus we can interpret this pattern as Brownian motion with drift in which an ancestry which finds itself on the wrong side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct'' side of the hybrid zone.  In other words, selected alleles that have, by chance, encroached deep into the other side of the hybrid zone have likely done so quite recently, as they cannot survive long in such hostile territory.  
\alisa{As a result, the haplotype that the selected locus resides on has few opportunities for recombination on the `wrong' side, which is reflected in long blocks of contiguous ancestry.}
%The recently of this movement results in few opportunities for recombination and is therefore reflected in long blocks of contiguous ancestry.  
\plr{This next stuff is said later, under "Assumptions"; maybe should be moved there.}
\yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms. This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.}

\subsection*{Additional signatures around selected loci}

\plr{(selection increases drift and migrant families are small; need better words/description here.)}
We further found that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will likely come from a relatively large number of independent ancestors per individual of that ancestry. (Figure~\ref{Fig:family_size}). % is this right 
This result also reflects the fact that, on average, haplotypes on the ``wrong'' side of the hybrid zone tend to be recent migrants and have therefore had few descendants.

\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding. We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}). Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short. A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site. Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will largely recombine with the alternate ancestry, and therefore decay rapidly. \yb{This finding suggests that, despite the fact that haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.}

\plr{(we'll be able to see this soon)}
These two novel findings may be beyond detection in current research programs that often make use of low-density genetic markers for which phase is unknown. However, as the const of sequencing continue to drop and long-phased reads become more common, access to data pertaining to these observations could further aid in identifying selected sites in hybrid zones.  %\yb{(CITE doi:10.1038/nbt0515-433 )} %Startups use short-read data to expand long-read sequencing market Michael Eisenstein Nature Biotechnology 33, 433?435 (2015) doi:10.1038/nbt0515-433 Published online 12 May 2015
\alisa{In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available data, surrounding selected loci could offer an approximation to these findings.}

\subsection*{Paths to inference}

\plr{(combine with haplotype length stuff above)}
Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $l_z(k\pm)$, could provide a simple and informative summary. 
This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.

@@

\plr{(het. disadvantage)}
We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations. %Thus before our results can be profitably used for inference, it is worthwhile to  reflect on our assumptions and their consequences. 
Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus. This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met. \alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}

\plr{(one locus only)}
The assumption about the density of selected loci, however, may have greater consequences. Our model, which focuses on a single selected site is likely to be still relevant to cases in which there are only a small number of selected loci scattered throughout the genome. However, our predictions may differ significantly from scenarios in which the density of selected sites is high. Specifically, numerous selected sites within one ancestry block would likely act to both  increase the strength of selection, and to generate a longer unit which cannot be effectively broken up by recombination   \citep{Barton1983,Kruuk1999}. Such regions would likely be surrounded by even greater chunks of unbroken ancestry -- a result reflecting  greater $s$ and relatively short map distance between linked incompatibilities -- \alisa{and will impact expectations of genome-wide patterns of clines and block lengths. In terms of detecting targets of selection, such sites will be a common limitation among many approaches to identifying selection.}

>>>>

%\usepackage[hidelinks]{hyperref}

@@

The process of speciation commonly involves allopatric divergence in relative isolation \citep{Coyne2004}.  
When formerly isolated populations come back into contact 
before reproductive isolation is complete, gene flow is still possible. 
As two such populations interbreed,
individuals' genomes can be labeled according to which ancestral population they are inherited from,
and local interbreeding and migration of such populations
creates a gradient of ancestry assignments across geography,
centered on their point of contact \citep{Barton1985}. 
\plr{(check that neutral clines discussed in barton '85)}\alisa{This is simply stated in Barton1985 without any justification or citation. Can we just cite Sedghifar2015?}
If the populations are sufficiently diverged,
this process leaves a distinct pattern of variation across the genome, 
in which long tracts of divergent haplotypes from each ancestral population  
are broken up by historical recombination events,
leading to substantial correlations in ancestry between distant markers.

@@

Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center. For this, we build on the neutral framework presented in \citet{Sedghifar2015} to derive the expected distribution of ancestry block lengths across a contact zone.  
\yb{The theoretical expectations developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}

%\plr{explain relationship to \citet{Sedghifar2015}}

@@

\paragraph{Establishment of the cline at the selected locus}
The deterministic theory predicts that after secondary contact,
the two alleles at the selected locus will form a stable cline \citep{Barton1979},
affecting neighboring loci as well.
First, we turn our attention to how the cline at the selected locus itself is formed.
Let $p(x,t)$ denote the frequency of the $A$ allele at spatial location $x$ and time $t$.  
Suppose that secondary contact occurred at time $t=0$, 
and that $p(x,0) = 1$ if $x<0$ and $p(x,0)=0$ if $x>0$.
% (In discrete computations, a deme at $x=0$ has $p(0,0)=1/2$.)
Let $\sigma^2$ be the mean squared distance between parent and offspring 
(i.e., the variance of the distribution of dispersal distance).
Assuming that: (i) alleles locally assort into diploids randomly, 
(ii) habitat and dispersal are homogeneous, and 
(iii) $s$ and $\sigma$ are small, then the commonly-used equation that $p$ solves is
\begin{align} \label{eqn:cline_pde}
    \dot p = \frac{\sigma^2}{2} \Delta p + s p (1-p) (2p-1) ,
\end{align}
where $\dot p(t,x)$ is the time derivative and $\Delta p$ is the Laplacian,
as derived in \citet{Bazykin1969}; for discussion of diploidy see \citep{christiansen1995genotypic}.
Although there is no known exact analytic solution to this equation, even in steady state, 
a good approximation to the equilibrium solution is that
$\lim_{t \to \infty} p(x,t) \approx (1+\tanh(-2x\sqrt{s}/\sigma))/2$ \citep{Bazykin1969}. 
This relies on numerous approximations, but the general conclusions should be quite robust: 
the stable cline has width of order $\sigma/\sqrt{s}$, and decays exponentially. 
Note that rescaling space and time by $\sigma/s$ and $1/s$ respectively in equation \eqref{eqn:cline_pde} results in a dimensionless equation,
implying that the cline establishes over a timescale of the order $1/s$. 
In other words, the cline is established through diffusion, which is slowed down somewhat by selection against heterozygotes. 
%\alisa{I don't think I understand the independence from $\sigma$?}
%\plr{It gets set up at speed $\sigma$ over a distance proportional to $\sigma$.}

@@

At a geographic position far from the center, $p=1$ to the left, and is proportional to  $\exp(-x\sqrt{s}/\sigma)$ to the right.
Therefore, roughly, lineages on ``their own'' side wander randomly,
while lineages on ``the wrong'' side are pushed at constant speed $\sigma/\sqrt{s}$ 
back towards the side where they are more common. 
Since an $A$ allele must by definition have been inherited from the $A$ side of the barrier 
at the time of secondary contact, this push must get more intense the closer it is to the time of secondary contact.
Lineages of $B$ alleles do the same thing, of course, but in the opposite direction. 
% \alisa{I want to make a figure from the simulations that show the rapid retreat of wrong ancestries back to where they came from}

@@

\paragraph{PDE for linked clines}
We can use this diffusion model for lineages to compute expected clines in ancestry,
i.e., the expected proportion of individuals who inherit from species $A$,
as a function of space, time, and position on the genome.
Precisely, we need the probability that 
an allele sampled $t$ generations after secondary contact at location $x$,
at recombination distance $r$ to a selected allele of type $z$,
is inherited from an individual of ancestry $A$,
where $z$ can be either $A$ or $B$.
We denote this probability $q_z(x,t,r)$.
In other words,
\begin{align}
    q_z(x,t,r) = \P^x \{Z_t = A\} .
\end{align}
(Recall that $Z$ depends implicitly on $r$.)
The description above implies that $q_z$ solves the following Kolmogorov backward equation:
\begin{align}
    \begin{aligned}  \label{eqn:q_pde}
    \dot q_A(x,t,r) 
            &= \sigma^2 \grad \log(p(x,t)) \cdot \grad q_A(x,t,r) 
                + \frac{\sigma^2}{2} \Delta q_A(x,t,r) 
            \\ &\qquad {} + 
                r (1-p(x,t))(q_B(x,t,r)-q_A(x,t,r))  \\
    \dot q_B(x,t,r) &= \sigma^2 \grad \log(1-p(x,t)) \cdot \grad q_B(x,t,r) 
            + \frac{\sigma^2}{2} \Delta q_B(x,t,r)
            \\ &\qquad {} + 
            r p(x,t) (q_A(x,t,r)-q_B(x,t,r))  ,
    \end{aligned} 
\end{align}
with boundary conditions $q_A(x,0,r)=1$ and $q_B(x,0,r)=0$, 
and where $\grad$ is the gradient, and $\Delta$ is the Laplacian.
%\alisa{this is fwd-in-time notation. should we change it to bwd-in-time?}
%\plr{This equation should be forwards in time, since it's the Kolmogorov backward equation
%applied to a backwards-time process!}

@@

Hybrid zones of varying ages were simulated under conditions of $\sigma=1$ and $s=0.1$ and $s=0.01$. From this we measure the length of unbroken tracts of ancestry $B$ surrounding a given position along a chromosome (if the locus in question was of ancestry $A$ the ancestry block had length $0$). These were compared to the output of neutral simulations with $s=0$.

The simulations were executed in R, with scripts available on
\url{https://github.com/asedghifar/hybridzones}.

@@

%Another statistic is the cumulative probability $C_I(m)$ of a sampled chunk encompassing position $m$ being longer than a chunk sampled from the distribution of all chunk lengths across the genome in the population.

Additionally, we look at the mean value of these statistics in the two chunks, $m-$ and $m+$, that flank the block of unbroken ancestry containing $m$. Since there appears to be a signal at these regions also, we further define a statistic

$$C =  \frac{2\sum{l_B(k_i)}}{\sum{l_A(k_i+)+l_A(k_i-)}}$$

which is the ratio of mean block length around $k_i$ and mean length of blocks directly flanking the block surrounding $k_i$.

@@

\begin{figure}
\includegraphics{figs/blocksAlongChromAncBConditioningHighRes}
\caption{Mean block length, $l_B(k_i)$, along a simulated chromosome of length $1M$ with selected locus at position $0.5M$. Here each line represents  $l_B(\mathbf{k})$, the mean block lengths in a given deme, and is normalized by mean block length $l_B(k_i)$ along the chromosome in the deme.
    \plr{what is $T$ and $s$? also, add version with different $T$ and/or $s$ to the supplement.}
}\label{Fig:blockLengths}
\end{figure}

@@

%\plr{This first paragraph sounds like the Introduction.  A brief recap here is good: maybe something like:
%    Our goal in this paper was to describe cline formation and haplotype structure around a new hybrid zone,
%   which we have done using both theory and simulation.
%   Our most useful observations are that clines establish over time $1/s$,
%    and that lineages of selected loci found on the side of the hybrid zone where they are at a disadvantage
%    tend to move back towards the side where they are common
%   at speed $\sigma/\sqrt{s}$.
%   This predicts that blocks of ancestry are longer near the center, flanked by shorter ones ...
%}

\alisa{Using a combination of theory and simulations, we present a description of the process of cline formation and haplotype structure in a relatively young (i.e. non-equilibrium) hybrid zone.
We show that clines establish over time $1/s$, and that lineages of selected loci tend to move back towards their `ancestral home' when in a geographic region where they  are unfit. 
This occurs at speed $\sigma/\sqrt{s}$. Based on this we predict, and observe in simulations, that blocks of ancestry surrounding these selected loci are longer, especially when distant from the center of the cline. 
This extends previous theoretical work on hybrid zones, which has primarily focused on stable clines in allele frequency and  suggests approaches to interpreting patterns of variation in hybrid zone populations, as well as detecting targets of purifying selection in hybrid zones.}

\subsection*{Genomic signals associated with targets of selection}

There is great interest in identifying the loci that are experiencing selection in hybrid zones. Popular approaches to this involve identifying alleles that are exceptional in terms of frequency across space, or genome-wide admixture proportion \citep{Porter1997,Gompert2012}.  
Current genomic resources, however, can provide an additional rich information when applied to the study of hybrid zones.
In particular, ancestry deconvolution facilitated by programs such as  hapMIX \citep{Price2009}, LAMP \citep{Sankararaman2008} and fineSTRUCTURE \citep{Lawson2012} provides researchers with a measure of the length of ancestry tracts across the genome and this information has been used extensively to infer the demographic history of hybridization/admixture from present day samples \citep[e.g. ][]{Hellenthal2014}. 
In order to interpret these patterns in the context of a hybrid zone, we set out to understand the impact of selection on patterns of ancestry tract lengths, and here present an expectation of block-length patterns throughout a hybrid zone. 
Importantly, we find that selection against hybrid incompatibilities leaves a distinct genomic signature in the form of long contiguous blocks of ancestry around these loci.
This pattern reflects the intuition that regions surrounding selected loci introgress less readily, and loci under selection can therefore be thought as moving under Brownian motion with drift whereby a lineage that is presently on the `` wrong" side of a hybrid zone rapidly moves, backwards in time, across space to the ``correct" side. 
In other words, disadvantageous alleles that have encroached deep into the other side of the hybrid zone have done so by chance and, since they do not persist for long on the ``wrong" side, are likely to have done so relatively recently. Haplotypes surrounding the selected locus therefore have had relatively few opportunities for recombination with different ancestries and this is reflected in longer blocks of contiguous ancestry.

Through our simulations we find also that blocks of ancestry that are have crossed the hybrid zone and are closely linked to the selected site without overlapping it (i.e., blocks that are adjacent to the block containing the selected site) are exceptionally short. This is measured by the length of the block adjacent to the block containing the selected site. 
A simple and intuitive explanation for this patter in that loci physically linked to a selected site can only survive on the wrong side of the hybrid zone if they recombine off the background of the unfit variant. 
A disproportionate majority of the subsequent recombination events it experiences will be with the more fit ancestry, and so the ancestry block will decay rapidly. 
This finding suggests that, despite the fact that  haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.

Our results suggest that a few natural statistical summaries could help identify selected loci in hybrid zones. 
Specifically, the statistic $l_z(k)$ -- the genetic length of ancestry surrounding a site of a given lineage -- and similarly,  and similarly $l_z(k\pm)$, could provide a simple and informative summary. 
This statistic, perhaps in relation to the length of ancestry around adjacent sites, could be developed into an approach to scan genomes from hybrid zones to identify putatively selected loci.

While these statistics presented here are currently detectable only in the few systems for which dense genetic markers and phase information is known, as the cost of sequencing continues to drop and long-phased reads become more common, the patterns described here could further aid in the identification of selected sites in hybrid zones. In the meantime, patterns of elevated pairwise LD, which is more easily computed from readily available low-density and unphased genomic data, could offer a less-powerful alternative to these findings.

@@

\alisa{These seem like an orphan statement. Any idea where to stick these?}
We further find that the long blocks of ancestry around selected sites on the wrong side of the hybrid zone will be descended from a relatively large number of independent ancestors, per individual, of the less fit ancestry. This result reflects that, on average, less fit haplotypes tend to be recent migrants with few descendants.

@@

We made numerous simplifying assumptions to derive our analytical and also in the execution of the simulations with respect to the architecture of selection, the geographic distribution of populations, and relatedness of individuals.

Although our assumed scenario of selection against heterozygotes at a single locus is uncommon in nature, we believe that this is unlikely to drastically influence our findings, as previous studies have demonstrated that underdominant loci share similar properties with more realistic models (e.g., ecological selection \citep{Barton1989,Barton1993}). 
We expect, therefore, our findings here to be somewhat generalizable.

Our assumption of a single selected locus may have greater consequences. While our model is likely to be relevant to cases in which there are only a small number of targets of selection scattered throughout the genome, our predictions may differ significantly from situations in which the density of selected sites is higher. 
Having numerous selected sites within one ancestry block increases the strength of selection, and the relatively short map distances between linked incompatibilities generates a longer unit which is not readily broken up by recombination. Both factors are expected to result in regions that are surrounded by even greater chunks of unbroken ancestry, and will impact expectations of genome-wide pattens of clines and block lengths, as well as the resolution to which one could detect targets of selection. 
 $s$ and relatively short map distances between linked incompatibilities)

%Foremost, we have assumed that selection acts against heterozygotes, and that there is a single underdominant locus. This situation is uncommon in nature, since underdominant loci cannot invade when rare, and thus this assumption is likely rarely met. \alisa{We believe, however, that this assumption is unlikely to drastically influence our results, and previous results demonstrating that underdominant loci share similar properties as more realistic genetic models such as ecological selection, mean that our findings here are somewhat generalizable} \citep{Barton1989, Barton1993}

\plr{This next stuff is said later, under "Assumptions"; maybe should be moved there.}
\yb{While our theory focusses on a spatially arranged hybrid zone, it is likely that our results extend to alternative models of hybridization including models of assortative mating and hybrid swarms. This is because regardless of the specific demographic model, selection against hybrids will severely limit opportunities for recombination between ancestries around incompatibilities.}

%\plr{(long blocks adjacent to short ones: combine with bit on long blocks)}
%In addition to our finding that ancestry surrounding selected sites is relatively long, our simulations revealed a consistent and unexpected finding. We found, on a given side of the hybrid zone, a transition from the locally adaptive ancestry at the selected site to the alternative ancestry at a closely linked site was short-lived (Figure~\ref{Supp:adjacentBlocks}). Put differently, blocks of `wrong' ancestry that do not overlap the selected site but are closely linked are exceptionally short. A simple intuitive explanation for this pattern suggests itself - loci physically linked to the selected sites can only survive on the wrong side of the hybrid zone for a reasonable amount of time if they recombine away from their ancestry at the locally adaptive site. Because it is then tightly linked to the alternate ancestry at the selected site, this ancestry chunk will largely recombine with the alternate ancestry, and therefore decay rapidly. \yb{This finding suggests that, despite the fact that haplotypes surrounding incompatibilities might be quite long, genome scans may have the power to extract fine-grained regions near selected loci from the large chunks of ancestry that will often flank these regions.}
